Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2019-04-09 article_springer/article_springer.Rmd
Change pictures, make everything reproducible

preamble: | 
  \usepackage{amssymb}
  \usepackage{natbib}
  \newtheorem{definition}{Definition}

@@

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("rstudio/rticles")
# use the newest MittagLeffleR version:
devtools::install_github("UNSW-MATH/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("UNSW-MATH/CTRE")
library(CTRE)
setwd("~/GitHub/bursty-POT/article_springer")
Sys.setenv(TEXINPUTS=getwd(),
           BIBINPUTS=getwd(),
           BSTINPUTS=getwd())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
set.seed(12345)
```

@@

\begin{definition}[MRP]
Let $(W,J), (W_1, J_1), (W_2, J_2), \ldots$ be i.i.d. pairs of random 
variables, where the $W_k > 0$ are interpreted as the \emph{waiting times}
and $J_k \in [x_L, x_R]$ as the \emph{event magnitudes}
($x_L \in [-\infty, +\infty), x_R \in (-\infty, +\infty]$). 
If $W$ and $J$ are independent, the Marked Renewal Process is said to be \emph{uncoupled}. 
\end{definition}

@@

\begin{definition}[CTRE]
Given a threshold $\ell \in (x_L, x_R)$, 
consider the stopping time 
$$\tau(\ell) := \min\{k: J_k > \ell\},\quad \ell \in (x_L, x_R).$$
Define the pair of random variables $(X(\ell), T(\ell))$ via 
$$X(\ell) = J_{\tau(\ell)} - \ell, \quad 
T(\ell) = \sum_{k=1}^{\tau(\ell)} W_k.$$
By restarting the MRP at $\tau(\ell)$, inductively define the
two i.i.d. sequences $T(\ell,n)$ and $X(\ell, n)$, $n \in \mathbb N$, 
called the "interarrival times" and the "exceedances", respectively.
The pair sequence $(T(\ell, n), W(\ell, n))_{n \in \mathbb N}$ is called 
a Continuous Time Random Exceedance model (CTRE). 
If the underlying MRP is uncoupled, then the CTRE is also called 
uncoupled. 
\end{definition}

```{r thresholdedBursty, message=FALSE, fig.height=3, fig.width=8, fig.cap="\\label{fig:thresholdedBursty}Exceedances (red) and Times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)

times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
#times<-cumsum(rpareto(n,tail))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
par(mfrow = c(1,2)) 
p1<-plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
p2<-flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data")
library(gridExtra)
grid.arrange(p1, p2, nrow = 2)
```

@@

Historically, the first method proposed for the estimation of the Mittag-Leffler distribution parameters was the fractional moment estimator by @kozubowski2001. Unlike the first moments, the fractional moments of order $p$ for $p<\beta$ exist and are tractable. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. @Cahoy2010 proposed a moment estimator of the log transformed data, which does not require any prior. Furthermore, they performed simulation studies illustrating that the log-Moment outperforms the fractional moment estimator with respect to bias and root mean squared error (RMSE).

@@

```{r MSE, echo=FALSE, fig.height=12, fig.cap="\\label{fig:MSE} MSE for sample sizes n=30, 100, 200 for the tail and scale estimations via log-moments and maximum likelihood estimation", out.width = '100%'}
knitr::include_graphics("MSE_n25_100_200.pdf")
```

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator
\begin{align*}
H_{k,n}= \frac{1}{k} \sum_{i=0}^{k-1} \log \frac{X_{n-i,n}}{X_{n-k,n}}. 
\end{align*}
However, these methods are statistically less efficient since they only use a portion of the information contained in the data, as mentioned by @kozubowski2001. Moreover, the hill estimator requires a tuning parameter, and only performs reliably well for distributions close to Pareto [see @Resnick97]. Figure \ref{fig:Hillplots} shows Hill plots for Mittag-Leffler simulated data, with varying sample sizes and tails. To deduce the correct tail parameter estimates from these plots is virtually impossible.

```{r Hillplots, echo=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hillplots for simulated Mittag-Leffler distributed rv's with tail parameter 0.5 (two plots in the right column) and 0.8 (two plots in the left column) and sample sizes 200 (two plots in the lower row) and 1000 (two plots in the upper row). ", out.width = '90%'}
 knitr::include_graphics("Hillplots.pdf")
```

Since the exponential distribution is nested in the Mittag-Leffler family of distributions, an appropriate way to choose between a model with exponential and Mittag-Leffler inter-exceedance times seems to be a Likelihood ratio test. Although the two models are nested, the assymptotic distribution is not $\chi^2_1$-distributed, the classical Theorem of Wilk doesn't apply since under $H_0$ the parameter $\beta$ of the Mittag-Leffler distribution is equal to one and hence lies under $H_0$ on the boundary of the parameter space $(0,1]$. Nonetheless, one can perform a bootstrapped log-likelihood ratio test. In Figure \ref{Fig:LRT} one can see the power for the bootstrapped LRT for Mittag-Leffler distributions with different tail parameters. As expected the power gets smaller for tail parameters close to one, since the Mittag-Leffler distribution converges for $\beta \rightarrow 1$ to an exponential distribution. Consequently, for tail parameters close to one it is hard to differentiate a Mittag-Lefller distribution from an exponential.

```{r likelihood-ratio-test}
flares_arrivals <- flares %>% ctre() %>% thin(k = 500) %>% interarrival()
mlmle_out <- mlmle(data = flares_arrivals)
```

Since the exponential distribution is nested in the Mittag-Leffler family of 
distributions, a Likelihood Ratio Test seems to be an appropriate way to choose 
between a model with exponential and Mittag-Leffler inter-exceedance times.
Although the two models are nested, the asymptotic distribution is not 
$\chi^2_1$-distributed, and Wilk's Theorem does not hold: under $H_0$, the 
parameter $\beta$ of Mittag-Leffler distribution is equal to $1$, and hence lies
on the boundary of the parameter space $(0,1]$. 
Instead, a valid approach is a bootstrapped log-likelihood ratio test 
(PUT REFERENCE HERE). 
Figure \ref{Fig:LRT} displays the power for the bootstrapped LRT for Mittag-Leffler distributions with varying tail parameters. As expected, the power decreases for tail parameters close to one, since the Mittag-Leffler distribution converges as $\beta \uparrow 1$ to an exponential distribution; it becomes hard to differentiate a Mittag-Lefller distribution from an exponential. 
(PUT ALL 4 CURVES INTO ONE PLOT)

```{r LRT power, echo=FALSE, fig.align="center", fig.cap="\\label{Fig:LRT} Power for bootstrapped LRT for different sample sizes", out.width = '80%'}
knitr::include_graphics("Plots_LRT_boot.pdf")
```

@@

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the fitted threshold exceedance time distribution is 
$$
T(\ell) \sim {\rm ML}(\beta, k^{-1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not memoryless, and the probability density of the time $t$ until the next threshold crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{-1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{-1/\beta} \sigma_0)$ is the probability density of ${\rm ML}(\beta, -k^{1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{-1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{-1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a decreasing function for the Mittag-Leffler distribution. 
```{r hazard, fig.height=4, fig.cap="Left: Conditional distribution of time until next threshold crossing, depending on elapsed time $t_0$ since last crossing ($\\beta = 0.8$, $\\sigma_0 = 1$). Right: Hazard rate depending on tail parameter $\\beta$.\\label{fig:hazard}"}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(1,2))
plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
for (k in 1:length(t_0)) {
  cond <- pml(q = t_0[k], tail = tail, scale = scale, lower.tail = FALSE)
  yy <- dml(x = xx + t_0[k], tail = tail, scale = scale) / cond
  lines(xx,yy, col = k+1, lty = k)
}
legend("bottomleft", c("t_0 = 0", "t_0 = 1", "t_0 = 10"), col = 2:4, lty = 1:3)

betas <- c(0.7, 0.8, 0.95, 0.99)
t_0 <- 1
plot(xx, rep(1, length(xx)), ylim = c(0.001,2), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'h')
for (k in 1:length(betas)) {
  yy <- dml(x = xx, tail = betas[k], scale = scale) / pml(
    q = xx,
    tail = betas[k],
    scale = scale,
    lower.tail = FALSE
  )
  lines(xx,yy, col = k+1, lty = k)
}
legend("bottomleft", c("beta = 0.7", "beta = 0.8", "beta = 0.95", "beta = 0.99"), col = 2:5, lty = 1:4)

@@

To test our inference method via stability plots, we have simulated 
independent waiting time and magnitude pairs $(W_k, J_k)$ for three different
situations: stable waiting times, Pareto waitings times and Exponential waiting 
times. In order to have exact 
analytical values available for $\beta$ and $\sigma_0$, a distribution for 
$W_k$ needs to be chosen for which $b(n)$ from \eqref{eq:stability} is known. 
If we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, 
then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation [see e.g. the vignette on parametrisation in @MittagLeffleR]
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist.

@@

<!-- ```{r simulated-example, message=FALSE, warning=FALSE, fig.cap=caption} -->
<!-- k_max <- 500 -->

<!-- thin_sim_ctre <- thin(ctre = sim_ctre, k = k_max) -->
<!-- # par(mfrow=c(1,2)) -->
<!-- MLestimates(ctre = thin_sim_ctre, tail = params$tail, scale = n^(1/params$tail)) -->
<!-- ``` -->

By the Theorem, the distribution 
of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$. we have chosen $\beta=1$ in the stable as well in the
Pareto case. Figure \ref{fig:Ex1figs} displays plots of $\hat \beta(k)$ vs. $k$ for
the first example with stable distributed waiting times. In the left picture we 
used the log moment estimator, in the middle picture maximum likelihood estimates 
(both applied to the inter-exceedance times) and for the right picture hill 
estimators applied to the inter-arrival times. Recall that $k$ is the index of the 
order statistics of $J_k$ at which the threshold $\ell$ is placed.

```{r Ex1figs, echo=FALSE,fig.height=7,fig.cap="\\label{fig:Ex1figs} Stability Plots for m=100 simulation runs for Example 1 (stable distributed waiting times, true tail parameter =0.8). The grey lines, the red line and the blue lines are the log Moment, the MLE and the hill estimator, repectively, for different $k$'s on the x-axis. The dark lines in each picture are the means of the concerning estimators for different $k$'s.", fig.show='hold', out.width = '32%'}
knitr::include_graphics(c("SimuStudy_StabilityPlots_Ex1_Tail08_logMoms.pdf","SimuStudy_StabilityPlots_Ex1_Tail08_mles.pdf","SimuStudy_StabilityPlots_Ex1_Tail08_hill.pdf"))
```
Figure \ref{fig:Ex2figs} shows the estimates for different $k$'s and for the
different estimators for the tail parameter $\beta$. Again one can see that 
the Hill estimator has a higher variance.

```{r Ex2figs, echo=FALSE,fig.height=7,fig.cap="\\label{fig:Ex2figs} Stability Plots for m=100 simulation runs for Example 2 (Pareto distributed waiting times, true tail parameter =0.8). The grey lines, the red line and the blue lines are the log Moment, the MLE and the hill estimator, repectively, for different $k$'s on the x-axis. The dark lines in each picture are the means of the concerning estimators for different $k$'s.", fig.show='hold', out.width = '32%'}
knitr::include_graphics(c("SimuStudy_StabilityPlots_Ex2_Tail08_logMoms.pdf","SimuStudy_StabilityPlots_Ex2_Tail08_mles.pdf","SimuStudy_StabilityPlots_Ex2_Tail08_hill.pdf"))
```
To show how the different estimates behave in the case where $\beta=1$
(the exponential case), we simulated in a last example exponentially distributed
waiting times.

```{r Ex3figs, echo=FALSE,fig.height=7,fig.cap="\\label{Fig:Ex3figs} Stability Plots for m=100 simulation runs for Example 3 (Exponentially distributed waiting times, true tail parameter =0.8). The grey lines, the red line and the blue lines are the log Moment, the MLE and the hill estimator, repectively, for different $k$'s on the x-axis. The dark lines in each picture are the means of the concerning estimators for different $k$'s.", fig.show='hold', out.width = '32%'}
knitr::include_graphics(c("SimuStudy_StabilityPlots_Ex3_logMoms.pdf","SimuStudy_StabilityPlots_Ex3_mles.pdf","SimuStudy_StabilityPlots_Ex3_hill_times.pdf"))
```

>>>>

preamble: | 
  \usepackage{amssymb}
  \usepackage{natbib}

@@

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("rstudio/rticles")
# use the newest MittagLeffleR version:
devtools::install_github("UNSW-MATH/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("strakaps/CTRE")
library("CTRE")
library(CTRE)
install.packages("here")
library(here)
Sys.setenv(TEXINPUTS=getwd(),
           BIBINPUTS=getwd(),
           BSTINPUTS=getwd())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
library(gridExtra)
library(ggplot2)
library(dplyr)
set.seed(12345)
```

@@

**Definition (MRP):**

: Let $(W,J), (W_1, J_1), (W_2, J_2), \ldots$ be i.i.d. pairs of random 
variables, where the $W_k > 0$ are interpreted as the *waiting times* 
and $J_k \in [x_L, x_R]$ as the *event magnitudes* 
($x_L \in [-\infty, +\infty), x_R \in (-\infty, +\infty]$). 
If $W$ and $J$ are independent, the Marked Renewal Process is said to be *uncoupled*.

@@

**Definition (CTRE):**

: Given a threshold $\ell \in (x_L, x_R)$, 
consider the stopping time 
$$\tau(\ell) := \min\{k: J_k > \ell\},\quad \ell \in (x_L, x_R).$$
Define the pair of random variables $(X(\ell), T(\ell))$ via 
$$X(\ell) = J_{\tau(\ell)} - \ell, \quad 
T(\ell) = \sum_{k=1}^{\tau(\ell)} W_k.$$
By restarting the MRP at $\tau(\ell)$, inductively define the
two i.i.d. sequences $T(\ell,n)$ and $X(\ell, n)$, $n \in \mathbb N$, 
called the "inter-arrival times" and the "exceedances", respectively.
The pair sequence $(T(\ell, n), W(\ell, n))_{n \in \mathbb N}$ is called 
a Continuous Time Random Exceedance model (CTRE). 
If the underlying MRP is uncoupled, then the CTRE is also called 
uncoupled.

```{r thresholdedBursty, message=FALSE, fig.height=6, fig.width=7,out.width='70%',fig.align='center',fig.cap="\\label{fig:thresholdedBursty}Exceedances (red) and times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)
times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
p1<-plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
p2<-flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
grid.arrange(p1,p2)

```

@@

Historically, the first method proposed for the estimation of the Mittag-Leffler distribution parameters was the fractional moment estimator by @kozubowski2001. Unlike the first moments, the fractional moments of order $p$ for $p<\beta$ exist and are tractable. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. @Cahoy2010 proposed a moment estimator of the log-transformed data, which does not require any prior. Furthermore, they performed simulation studies illustrating that the log-moment outperforms the fractional moment estimator with respect to bias and root mean squared error (RMSE).

@@

```{r MSE, echo=FALSE, fig.height=7,fig.width=8,fig.align="center", fig.cap="\\label{fig:MSE}MSE for sample sizes n=25, 100, 200 for tail and scale estimation via log-moments and mle", out.width = '90%'}
load(here::here("article_springer\\article_springer_files\\ML_estim_simu","tbl_ges.RData"))
tbl_ges %>% 
  filter(n == "25" || n == "100" || n=="200") %>%
  filter(true.scale == "1") %>% 
  ggplot(aes(true.tail, mse, colour = method)) + 
  geom_line() + geom_point() + 
  facet_wrap(c("n","param"), scales = "free", ncol = 2) + 
  scale_color_manual(values=c("red", "blue"))
```

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple]
\begin{align*}
H_{k,n}= \frac{1}{k} \sum_{i=0}^{k-1} \log \frac{X_{n-i,n}}{X_{n-k,n}}. 
\end{align*}
However, these methods are statistically less efficient since they only use a portion of the information contained in the data, as mentioned by @kozubowski2001. Moreover, the hill estimator requires a tuning parameter, and only performs reliably well for distributions close to Pareto [see @Resnick97]. Figure \ref{fig:Hillplots} shows Hill plots for Mittag-Leffler simulated data, with varying sample sizes and tails. To deduce the correct tail parameter estimates from these plots is virtually impossible.

```{r Hillplots, echo=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots}Hillplots for simulated Mittag-Leffler data with sample size 200 (left) and 1000 (right). ", out.width = '90%'}
hill_estim<-function(k,data){
  n<-length(data)
  data<-sort(data)
  data_thinned<-data[(n-k):n]
  r<-1/k * sum(log(data_thinned/data_thinned[1]))
  return(1/r)
}

hill_estim_plot<-function(k_min,data,k_max=n,true_tail){
  n<-length(data)
  x_values<-k_min:k_max
  y_values<-sapply(x_values,hill_estim,data=data)
  df<-data.frame(x=x_values,y=y_values)
  ggplot2::ggplot(data=df,aes(x_values,y_values))+ geom_line(color="blue")+xlab("Order Statistics") + ylab("Hill-Estimator of Tail Index") +
  coord_cartesian(ylim=c(0,1.5)) + geom_hline(yintercept=true_tail, linetype="dashed", color = "red", size=0.5)
}

data_ml_1<-rml(n=200,tail=0.8,scale=1)
p1<-hill_estim_plot(k_min=30,data=data_ml_1,true_tail=0.8)

data_ml_2<-rml(n=1000,tail=0.8,scale=1)
p2<-hill_estim_plot(k_min=30,data=data_ml_2,true_tail=0.8)
grid.arrange(p1,p2,ncol=2)
```

Since the exponential distribution is nested in the Mittag-Leffler family of 
distributions, a Likelihood Ratio Test seems to be an appropriate way to choose 
between a model with exponential and Mittag-Leffler inter-exceedance times.
Although the two models are nested, the asymptotic distribution is not 
$\chi^2_1$-distributed, and Wilk's Theorem does not hold: under $H_0$, the 
parameter $\beta$ of Mittag-Leffler distribution is equal to $1$, and hence lies
on the boundary of the parameter space $(0,1]$. 
Instead, a valid approach is a bootstrapped log-likelihood ratio test 
[see e.g. @davison1997bootstrap]. 
Figure \ref{Fig:LRT} displays the power for the bootstrapped LRT for Mittag-Leffler distributions with varying tail parameters. As expected, the power decreases for tail parameters close to one, since the Mittag-Leffler distribution converges as $\beta \uparrow 1$ to an exponential distribution; it becomes hard to differentiate a Mittag-Lefller distribution from an exponential.

```{r LRT_power, echo=FALSE, fig.align="center", fig.cap="\\label{Fig:LRT} Power for bootstrapped LRT for different sample sizes",out.width='90%'}
load("C:/Users/hees/Dropbox (Uni)/UniDortmund/Forschung/ForschungCTRMs/Bursty_Pot_Paper_Simus/Simulationstudy_LRT_boot/estimates_df_LRT_boot1_01.RData")
df_LRT_boot_01 %>% filter(n=="20" || n=="50" || n=="100" || n=="200") %>% ggplot(aes(x=true.tail,y=power)) +  geom_line(aes(y=power,group=n,color=factor(n)),linetype="solid",size=0.7)+
  labs(y="Power",x= "True tail parameter")+
  scale_color_manual(values=c("red", "blue","darkgreen","black"),name  ="Sample size")
```

@@

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the fitted threshold exceedance time distribution is 
$$
T(\ell) \sim {\rm ML}(\beta, k^{-1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not memoryless, and the probability density of the time $t$ until the next threshold crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{-1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{-1/\beta} \sigma_0)$ is the probability density of ${\rm ML}(\beta, -k^{1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{-1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{-1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a decreasing function for the Mittag-Leffler distribution. 
```{r hazard, fig.height=4, fig.cap="Left: Conditional distribution of time until next threshold crossing, depending on elapsed time $t_0$ since last crossing ($\\beta = 0.8$, $\\sigma_0 = 1$). Right: Hazard rate depending on tail parameter $\\beta$.\\label{fig:hazard}"}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(1,2))
df1<-data.frame(x=xx,y=rep(1, length(xx)))
col<-c("red","blue","darkgreen","black")

plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
for (k in 1:length(t_0)) {
  cond <- pml(q = t_0[k], tail = tail, scale = scale, lower.tail = FALSE)
  yy <- dml(x = xx + t_0[k], tail = tail, scale = scale) / cond
  lines(xx,yy, col = col[k], lty = k+1)
}
legend("bottomleft", c("t_0 = 0", "t_0 = 1", "t_0 = 10"), col = c("red","blue","darkgreen"), lty = 2:4)

betas <- c(0.7, 0.8, 0.95, 0.99)
t_0 <- 1

plot(xx, rep(1, length(xx)), ylim = c(0.001,2), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'h')
for (k in 1:length(betas)) {
  yy <- dml(x = xx, tail = betas[k], scale = scale) / pml(
    q = xx,
    tail = betas[k],
    scale = scale,
    lower.tail = FALSE
  )
  lines(xx,yy, col = col[k], lty = k+1)
}
legend("bottomleft", c("beta = 0.7", "beta = 0.8", "beta = 0.95", "beta = 0.99"), col = c("red","blue","darkgreen","black"), lty = 2:5)

@@

To test our inference method via stability plots, we have simulated 
independent waiting time and magnitude pairs $(W_k, J_k)$ (see upper panel in 
Figure \ref{fig:thresholdedBursty}) for three different situations: stable waiting times, Pareto waitings times and Exponential waiting times. In order to have exact 
analytical values available for $\beta$ and $\sigma_0$, a distribution for 
$W_k$ needs to be chosen for which $b(n)$ from \eqref{eq:stability} is known. 
If we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, 
then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation [see e.g. the vignette on parametrisation in @MittagLeffleR]
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist.

@@

By the Theorem, the distribution 
of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$. we have chosen $\beta=1$ in the stable as well in the Pareto case. Figure \ref{Ex1figs} displays plots of $\hat \beta(k)$ vs. $k$ for the first example with stable distributed waiting times. In the left picture we used the log-moment estimator, in the middle picture maximum likelihood estimates (both applied to the inter-exceedance times) and for the right picture hill estimators applied to the inter-arrival times. Recall that $k$ is the index of the order statistics of $J_k$ at which the threshold $\ell$ is placed.

```{r Ex1figs, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:Ex1figs} Stability Plots for m=100 simulation runs for Example 1 (stable distributed waiting times, true tail parameter =0.8). The grey lines, the red line and the blue lines are the log-moment, the MLE and the hill estimator, repectively, for different $k$'s on the x-axis. The dark lines in each picture are the means of the concerning estimators for different $k$'s.", fig.show='hold', out.width = '90%'}

path<-here::here("article_springer/article_springer_files/Simulation_results")
files<-list.files(path)
file_dirs<-paste(path,"/",files,sep="")
invisible(sapply(file_dirs,FUN=load,envir=.GlobalEnv,verbose=FALSE))

#Tail Parameter

# Example One

#logMoment estimator
logMom_Ex1_tail<-ggplot()+ geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_tail08_logMoms_Ex1) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_logMoms_Ex1) + ylab("Estimated tail") + geom_hline(yintercept=0.8, linetype="dashed", color = "red")

#ml estimator
mle_Ex1_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_tail08_mles_Ex1) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_mles_Ex1) + ylab("Estimated tail") + geom_hline(yintercept=0.8, linetype="dashed", color = "red")

#Hill estimator
hill_Ex1_tail<-ggplot()+ geom_line(data=tbl_estimates_tail08_hill_Ex1,aes(x=k,y=value,group=rep),colour="lightgrey") +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(data=means.tbl_hill_Ex1,aes(x=k, y=mw),colour="black") + ylab("Estimated tail") +  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

#Example Two

#logMoment estimator
logMom_Ex2_tail<-ggplot()+ geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Ex2_Pareto_tail08_logMoms) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_logMoms_Ex2) + ylab("Estimated tail") + geom_hline(yintercept=0.8, linetype="dashed", color = "red")

#ml estimator
mle_Ex2_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Ex2_Pareto_tail08_mles) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_mles_Ex2) + ylab("Estimated tail") + geom_hline(yintercept=0.8, linetype="dashed", color = "red")

#Hill estimator
hill_Ex2_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Ex2_Pareto_tail08_hill_times) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex2) + ylab("Estimated tail") +  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

#Example Three

#logMoment estimator
logMom_Ex3_tail<-ggplot()+ geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Example3_Exp_logMoms) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_logMoms_Ex3) + ylab("Estimated tail") + geom_hline(yintercept=1, linetype="dashed", color = "red")

#ml estimator
mle_Ex3_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Example3_Exp_mles) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_mles_Ex3) + ylab("Estimated tail") + geom_hline(yintercept=1, linetype="dashed", color = "red")

#Hill estimator
hill_Ex3_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Example3_Exp_hill_times) +
  coord_cartesian(ylim = c(0.5, 10)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex3) + ylab("Estimated tail") + geom_hline(yintercept=1, linetype="dashed", color = "red")

grid.arrange(logMom_Ex1_tail,mle_Ex1_tail,hill_Ex1_tail,logMom_Ex2_tail,mle_Ex2_tail,hill_Ex2_tail,logMom_Ex3_tail,mle_Ex3_tail,hill_Ex3_tail,ncol=3)

```
Figure \ref{Ex2figs} shows the estimates fordifferent $k$'s and for the different estimators for the tail parameter $\beta$. Again one can see, that there is much more variance between the stability plots for the different simulation runs in case of the hill estimator compared to the other two estimators.To show how the different estimates behave in the case that $\beta=1$, hence, the inter-exceedance times are exponetial distributed, we simulated in a last example exponetial distributed waiting times.

```{r Ex1figsScale, echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:Ex1figsScale} Stability Plots for the scale parameter for Example 1 (stable distributed waiting times, true scale parameter = 1).", fig.show='hold', out.width = '90%'}
#Scale Parameter

#Example One
logMom_Ex1_scale<-ggplot()+ geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_scales_logMoms_Ex1) +
  coord_cartesian(ylim = c(0, 2)) +
  geom_line(aes(x=k, y=mw),colour="black",means.tbl_scales_logMoms_Ex1) + ylab("Estimated scale") + geom_hline(yintercept=1, linetype="dashed", color = "red")

mle_Ex1_scale<-ggplot()+ geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_scales_mles_Ex1) +
  coord_cartesian(ylim = c(0, 2)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_scales_mles_Ex1) + ylab("Estimated scale") + geom_hline(yintercept=1, linetype="dashed", color = "red")

#Example Two
